Meeting the Challenges of Modeling Astrophysical Thermonuclear Explosions: \castro, \maestro, and the \amrex\ Astrophysics Suite

### 2017-11-01 astronum_2017/astronum-amrex.tex
typo
Added my middle initial.
righthand size -> righthand side.

\author{M. Zingale$^1$,
        A.~S. Almgren$^2$,
        M.~G. Barrios Sazo$^1$,
        V.~E. Beckner$^2$,
        J.~B. Bell$^2$,
        B. Friesen$^{3,2}$,
        A.~M. Jacobs$^4$,
        M.~P. Katz$^5$,
        C.~M. Malone$^6$,
        A.~J. Nonaka$^2$,
        D. Willcox$^1$, and
        W. Zhang$^2$}

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.48\linewidth]{summitdev_scaling}
\includegraphics[width=0.48\linewidth]{cori_scaling}
\caption{\label{fig:knl-gpu-scaling} \castro\ scaling on new
  architectures.  The left figure shows the \castro\ GPU proxy app
  (\starlord) on OLCF Summitdev.  We see good scaling behavior to 64
  GPUs in this test of a pure hydro problem (Sedov w/ real EOS).  On the
right, we explore performance on NERSC Cori (and Intel Xeon Phi platform).}
\end{figure}

@@

A parallel effort is porting our microphysics---in particular
reaction networks---to GPUs.  Our strategy is to do the entire ODE
integration on the GPU, i.e., the data for a patch of zones is passed
to the GPU, all righthand size and Jacobian evaluations and the
timestepping itself is done on the GPU, and once the burning in all
zones is completed, we access the data as needed on the CPU.  To
enable this, we ported our workhorse ODE integrator (VODE~\cite{vode})
to CUDA Fortran. This required extensive rewrites of the internals of
VODE, which was originally written using Fortran 77 syntax unsupported
by CUDA Fortran. Accelerating VODE with CUDA Fortran has proven
successful, and we now see significant performance gains with the CUDA
version of our reaction networks, even for moderate-sized networks (a
13-isotope standard network).  Figure~\ref{fig:cudaode} shows the
speed-ups on a GPU vs.\ single CPU core.  The main issue with scaling
is running out of local stack memory per thread with larger networks.
Reducing the memory footprint is a near-term goal for this work.  This
test problem also shows a lot of thread divergence due to the widely
differing thermodynamic conditions in the zones that are burning.

>>>>

\author{M. Zingale$^1$,
        A.~S. Almgren$^2$,
        M.~G. Barrios Sazo$^1$,
        V.~E. Beckner$^2$,
        J.~B. Bell$^2$,
        B. Friesen$^{3,2}$,
        A.~M. Jacobs$^4$,
        M.~P. Katz$^5$,
        C.~M. Malone$^6$,
        A.~J. Nonaka$^2$,
        D.~E. Willcox$^1$, and
        W. Zhang$^2$}

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.48\linewidth]{summitdev_scaling}
\includegraphics[width=0.48\linewidth]{cori_scaling}
\caption{\label{fig:knl-gpu-scaling} \castro\ scaling on new
  architectures.  The left figure shows the \castro\ GPU proxy app
  (\starlord) on OLCF Summitdev.  We see good scaling behavior to 64
  GPUs in this test of a pure hydro problem (Sedov w/ real EOS).  On the
right, we explore performance on NERSC Cori (an Intel Xeon Phi platform).}
\end{figure}

@@

A parallel effort is porting our microphysics---in particular
reaction networks---to GPUs.  Our strategy is to do the entire ODE
integration on the GPU, i.e., the data for a patch of zones is passed
to the GPU, all righthand side and Jacobian evaluations and the
timestepping itself is done on the GPU, and once the burning in all
zones is completed, we access the data as needed on the CPU.  To
enable this, we ported our workhorse ODE integrator (VODE~\cite{vode})
to CUDA Fortran. This required extensive rewrites of the internals of
VODE, which was originally written using Fortran 77 syntax unsupported
by CUDA Fortran. Accelerating VODE with CUDA Fortran has proven
successful, and we now see significant performance gains with the CUDA
version of our reaction networks, even for moderate-sized networks (a
13-isotope standard network).  Figure~\ref{fig:cudaode} shows the
speed-ups on a GPU vs.\ single CPU core.  The main issue with scaling
is running out of local stack memory per thread with larger networks.
Reducing the memory footprint is a near-term goal for this work.  This
test problem also shows a lot of thread divergence due to the widely
differing thermodynamic conditions in the zones that are burning.
